# Reproducible Research: Peer Assessment 1
```{r}
library(lattice)
```
## Loading and preprocessing the data
Load csv data **activity.csv** and convert dates to **R Date class format**
```{r}
actdata <- read.csv("activity.csv")
actdata$date <- as.Date(actdata$date,"%Y-%m-%d")
head(actdata)
```
Preprocess: Group days and sum steps.
```{r}
StepsSum <- tapply(actdata$steps, actdata$date,sum)
```

Preprocess: Group intervals and calculate mean of steps.
```{r}
StepsMean <- tapply(actdata$steps,actdata$interval,
mean,na.rm=TRUE)
```

## What is mean total number of steps taken per day?
1. Make a histogram of **total number of steps per day**
```{r}
hist(StepsSum,col="blue",xlab="Total Number of Steps Taken per Day",
ylab="Frequency", main="Histogram of Total Number of Steps taken per day")
```  

2. Calculate **mean** and **median** (ignoring the missing values).
```{r}
mean(StepsSum,na.rm=TRUE)
```
```{r}
median(StepsSum,na.rm=TRUE)
```
## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
plot(row.names(StepsMean),StepsMean,type="l",
xlab="Time Intervals (5-minute)",
ylab="Mean number of steps taken across all days",
main="Average Steps Taken at 5 minute Intervals",
col="blue")
```  

2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
i <- which.max(StepsMean)
i_max_steps <- names(i)
i_max_steps
```
The ** `r i_max_steps `** minute or ** `r i `th ** 5 minute interval reach the maximum number of steps  

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset.
```{r}
num_na_values <- sum(is.na(actdata))
num_na_values
```

2. Devise a strategy for filling in all of the missing values in the dataset.
```{r}
na_indices <- which(is.na(actdata))
imputed_values <- StepsMean[as.character(actdata[na_indices,3])]
names(imputed_values) <- na_indices
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled
Fill in missing values using the **average interval value across all days**
```{r}
for (i in na_indices) {
actdata$steps[i] = imputed_values[as.character(i)]
}
sum(is.na(actdata))
StepsTot <- tapply(actdata$steps, actdata$date,sum)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the **mean** and **median** total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
hist(StepsTot,col="red",xlab="Total Steps per Day",
ylab="Frequency", main="Histogram of Total Steps taken per day")
```
```{r}
mean(StepsTot,na.rm=TRUE)
```
```{r}
median(StepsTot,na.rm=TRUE)
```
The data is slighty impacted wiht of imputing missing data.

## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
days <- weekdays(actdata$date)
actdata$day_type <- ifelse(days == "Saturday" | days == "Sunday",
"Weekend", "Weekday")
StepsMean <- aggregate(actdata$steps,
by=list(actdata$interval,
actdata$day_type),mean)
names(StepsMean) <- c("interval","day_type","steps")
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
xyplot(steps~interval | day_type, StepsMean,type="l",
layout=c(1,2),xlab="Interval",ylab = "Number of steps")
tapply(StepsMean$steps,StepsMean$day_type,
function (x) { c(MINIMUM=min(x),MEAN=mean(x),
MEDIAN=median(x),MAXIMUM=max(x))})
```